---
title: "Group_Project"
format: html
editor: visual
---

## Quarto

## Libraries:

```{r}
library(haven)
library(dplyr)
library(ggplot2)
library(stringr) 
library(lubridate)
library(corrr)
library(ggplot2)
library(ggcorrplot)
library(tidyr)
library(scales)
library(mice)
library(readxl)
```

Data:

```{r}
#| echo: false
data <- read_dta("ZA7575.dta")
```

```{r}
#| echo: false
summary(data)
head(data)


column_info <- data.frame(
  column_name = colnames(data),
  column_label = sapply(data, function(x) attr(x, "label"))
)

column_info

data <- data %>%
  mutate(country_name = case_when(
    q1_1  == 1 ~ "Belgium",
    q1_2  == 1 ~ "Denmark",
    q1_3  == 1 ~ "Germany",
    q1_4  == 1 ~ "Greece",
    q1_5  == 1 ~ "Spain",
    q1_6  == 1 ~ "France",
    q1_7  == 1 ~ "Ireland",
    q1_8  == 1 ~ "Italy",
    q1_9  == 1 ~ "Luxembourg",
    q1_10 == 1 ~ "Netherlands",
    q1_11 == 1 ~ "Portugal",
    q1_12 == 1 ~ "United Kingdom",
    q1_13 == 1 ~ "Austria",
    q1_14 == 1 ~ "Sweden",
    q1_15 == 1 ~ "Finland",
    q1_16 == 1 ~ "Cyprus",
    q1_17 == 1 ~ "Czech Republic",
    q1_18 == 1 ~ "Estonia",
    q1_19 == 1 ~ "Hungary",
    q1_20 == 1 ~ "Latvia",
    q1_21 == 1 ~ "Lithuania",
    q1_22 == 1 ~ "Malta",
    q1_23 == 1 ~ "Poland",
    q1_24 == 1 ~ "Slovakia",
    q1_25 == 1 ~ "Slovenia",
    q1_26 == 1 ~ "Bulgaria",
    q1_27 == 1 ~ "Romania",
    q1_28 == 1 ~ "Croatia",
    q1_29 == 1 ~ "Other countries",
    q1_30 == 1 ~ "DK",
    TRUE ~ NA_character_  
  ))




data <- data |> 
  mutate(id = 1:nrow(data)) |>  # Create unique respondent_id
  select(id, everything())



```

## legal framework

QC19:Do you think that transgender persons should be able to change their civil documents to match their inner gender identity?(1=yes, 2=no,3=dk) QC20: Do you believe that official documents, like passports and birth certificates, should have a third option, such as X or O (other) beside male (M) and female (F)for those persons who do not identify as female and male? (1=yes, 2=no,3=dk) QC15_1: to what extend to do you agree or disagree with the following statement? Gay, lesbian and bisexual people should have the same rights as heterosexual people (1 2 3 4 5) QC15_2: to what extend to do you agree or disagree with the following statement? here is nothing wrong in a sexual relationship between two persons of the same sex(1 2 3 4 5) QC15_3: to what extend to do you agree or disagree with the following statement? Same sex marriages should be allowed throughout europe (1 2 3 4 5) QC4_7: In (OUR COUNTRY) when a company wants to hire someone and has the choice between two candidates with equal skills and qualifications, which of the following criteria may, in your opinion, put one candidate at a disadvantage? The candidate’s gender or sex (man or woman) QC4_8: In (OUR COUNTRY) when a company wants to hire someone and has the choice between two candidates with equal skills and qualifications, which of the following criteria may, in your opinion, put one candidate at a disadvantage? The candidate’s gender identity (being transgender) QC5_3: in the last 12 month have you done: You have joined an association or campaign that defends people against discrimination? (1=yes, 2=no,3=dk)

```{r}
#| echo: false

legal_framework_data <- data %>%
  select(
    qc19,        # TRANSGENDER - CHANGE CIVIL DOCUMENTS
    qc20,        # OFFICIAL DOCUMENTS SHOULD INCLUDE THIRD GENDER OPTION
    qc15_1,      # LGB STATEMENTS - SAME RIGHTS AS HETEROSEXUAL
    qc15_2,      # LGB STATEMENTS - NOTHING WRONG WITH SAME SEX RELATIONSHIPS
    qc15_3,      # LGB STATEMENTS - ALLOW SAME SEX MARRIAGE THROUGHOUT EUROPE
    qc4_7,       # JOB CANDIDATE DISADVANTAGE: GENDER
    qc4_8,       # JOB CANDIDATE DISADVANTAGE: GENDER IDENTITY
    qc5_3,       # ACTIONS AGAINST DISCRIMINATION - JOINED ASSOCIATION/CAMPAIGN
    country_name,
    id
  )

#convert to factor, for further processing 
legal_framework_data<- legal_framework_data %>%
  mutate(across(everything(), as.factor))


str(legal_framework_data)


desc_stats <- legal_framework_data %>%
  summarise(
    # Binary (0/1: y/n, 3 = idk)
    Mean_Binary_qc19 = mean(as.numeric(qc19), na.rm = TRUE),  
    Mean_Binary_qc20 = mean(as.numeric(qc20), na.rm = TRUE),  
    
    # Likert-Skalen (1-5)
    Mean_Likert_qc15_1 = mean(as.numeric(qc15_1), na.rm = TRUE),  
    Mean_Likert_qc15_2 = mean(as.numeric(qc15_2), na.rm = TRUE),
    Mean_Likert_qc15_3 = mean(as.numeric(qc15_3), na.rm = TRUE)
  )

print(desc_stats)

```

The analysis shows clear differences in support for various aspects of LGBT+ rights. It is particularly noticeable that both transgender-relevant topics and LGB rights tend to be rated positively overall, but with certain differences in the intensity of approval. The average agreement with the possibility of transgender people being able to adapt their official documents to their gender identity (qc19) is 1.59. This indicates that a majority of respondents are in favor of this measure, although “Don't Know” responses could play a role. The introduction of a third gender option on official documents (qc20) received slightly less approval with an average value of 1.72, which indicates that this measure is discussed somewhat more controversially.

In comparison, the values for the general equality of LGB people (qc15_1,, qc15_2 , qc15_3) are more on the approving side of the scale with average values between 2.17 and 2.38 (1 = strongly approve, 5 = strongly disapprove). This means that the majority of respondents tend to agree with these statements. However, the values are not extremely low, which indicates that there is still a certain skepticism or reluctance in parts of the population.

Identifiying how much people chooses the third option: "i dont know" in percentage:

```{r}
dk_percentage <- legal_framework_data %>%
  summarise(
    qc19 = sum(qc19 == 3, na.rm = TRUE) / n() * 100,
    qc20 = sum(qc20 == 3, na.rm = TRUE) / n() * 100
  )

dk_percentage

```

qc19 (Transgender people should be able to adapt official documents) → 11.95% qc20 (Third gender option in official documents) → 11.62%

Country differences:

```{r}

dk_summary <- legal_framework_data %>%
  group_by(country_name) %>%
  summarise(
    DK_qc19 = sum(qc19 == 3, na.rm = TRUE) / n() * 100,
    DK_qc20 = sum(qc20 == 3, na.rm = TRUE) / n() * 100
  ) %>%
  pivot_longer(cols = c(DK_qc19, DK_qc20), names_to = "Question", values_to = "DK_Percentage")


ggplot(dk_summary, aes(x = reorder(country_name, DK_Percentage), y = DK_Percentage, fill = Question)) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +
  coord_flip() +
  scale_fill_manual(values = c("DK_qc19" = "pink", "DK_qc20" = "orange"),
                    labels = c("qc19: Transgender documentchange", "qc20: third genderoption")) +
  labs(title = "Don't Know Rate ",
       x = "country", y = "% of dont know answers",
       fill = "Frage") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 10, face = "bold"))



```

-\> yyou can see clear differences in the response rates. While in bulgaria every 5. person selected "dont know" for transgender document change (qc19) in Belgium only less than 5% choosed this option.

Transgender Rights in Europe: Which Countries Support Legal Gender Changes?

```{r}
qc19_summary_sorted <- legal_framework_data %>%
  group_by(country_name) %>%
  mutate(total_responses = n()) %>%  
  group_by(country_name, qc19) %>%
  summarise(count = n(), total_responses = first(total_responses), .groups = "drop") %>%
  mutate(percentage = (count / total_responses) * 100) %>%
  filter(qc19 == "1") %>%
  arrange(percentage)


ggplot(qc19_summary_sorted, aes(x = reorder(country_name, percentage), y = percentage, fill = percentage)) +
  geom_bar(stat = "identity", color = "black", width = 0.7) +
  coord_flip() +
  scale_fill_gradientn(
    colors = c("red", "orange", "yellow", "green", "blue", "purple"), 
    values = rescale(c(min(qc19_summary_sorted$percentage), max(qc19_summary_sorted$percentage))),
    guide = "colorbar"
  ) +
  geom_text(aes(label = paste0(round(percentage, 1), "%")), hjust = -0.2, size = 5, fontface = "bold") +
  labs(
    title = "Support for Transgender Document Change (qc19) per Country",
    x = "Country",
    y = "Support Rate (%)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 16),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    legend.position = "right"
  )


```

Results: The Netherlands (88%), Spain (76%) and Malta (76%) are the countries with the highest approval ratings.Countries such as Germany (62%), Sweden (62%) and Austria (57%) also show a majority in favor of changing the law, but at a lower level than the frontrunners. In these countries, there appears to be ongoing social debate and differing opinions. Bulgaria (12%), Hungary (14%), Romania (22%) and Slovakia (24%) have the lowest approval ratings.

Because there arent so many variables for legal framework, we included the lgtb-rights index-data. The Data is from: https://ourworldindata.org/grapher/lgbt-rights-index?time=2012&region=Europe The index captures to which extent LGBT+ people have the same rights as straight and cisgender people. It combines 18 individual policies, such as the legality of same-sex relationships, marriage, and gender marker changes. Higher values indicate more rights, negative values regressive policies progressive policies receiving a positive score and regressive policies receiving a negative score: -5 to +13. Among the progressive policies are: Same-sex sexual acts Legal, Equal Age of Consent, Employment Discrimination, Hate Crime Protections, Incitement to Hatred, Civil Unions, Marriage Equality, oint Adoptions, Gender Marker Change, LGB Military, Transgender Military, Ban on Conversion Therapies, and Ban on Gender Assignment Surgeries on Children

```{r}
lgbt_index <- read.csv("lgbt-rights-index.csv")
str(lgbt_index)

lgbt_index <- lgbt_index |> filter(Year== 2019)

merged_data <- legal_framework_data %>%
  left_join(lgbt_index, by = c("country_name" = "Entity"))

ggplot(merged_data, aes(x = LGBT..Policy.Index, y = reorder(country_name, LGBT..Policy.Index), color = LGBT..Policy.Index)) +
  geom_point(size = 4) +
  scale_color_gradient(low = "red", high = "blue") +  
  labs(title = "LGBT Policy Index by Country",
       x = "LGBT Policy Index", y = "Country") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 10, face = "bold"))





```

Malta has the highest score, well above all other countries. Sweden, Netherlands, UK, Germany, France) have relatively high values. Eastern European countries (Latvia, Poland, Bulgaria, Romania, Hungary) have the lowest values in the LGBT Index.

We want to investigate whether there are significant differences in the laws in different parts of Europe. We therefore divided the countries into Eastern, Western, Southern and Northern Europe and compared the LGTB score.

```{r}
merged_data <- merged_data %>%
  mutate(region = case_when(
    country_name %in% c("Denmark", "Sweden", "Finland", "Estonia", "Latvia", "Lithuania") ~ "Northern Europe",
    country_name %in% c("Germany", "France", "Netherlands", "Belgium", "Luxembourg", 
                        "United Kingdom", "United Kingdom of Great Britain and Northern Ireland",  
                        "Ireland", "Austria") ~ "Western Europe",
    country_name %in% c("Spain", "Italy", "Portugal", "Greece", "Malta", "Cyprus") ~ "Southern Europe",
    country_name %in% c("Poland", "Hungary", "Czech Republic", "Slovakia", "Romania", 
                        "Bulgaria", "Croatia", "Slovenia") ~ "Eastern Europe",
    TRUE ~ "Other"
  ))


ggplot(merged_data, aes(x = region, y = LGBT..Policy.Index, fill = region)) +
  geom_boxplot() +
  labs(title = "LGBT Policy Index by Region",
       x = "Region", y = "LGBT Policy Index") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


             
```

Western Europe has the highest LGBT index, with a high median and a relatively small range. Northern Europe shows a large variance, with some countries having very high values and others in the middle. Southern Europe is lower overall than Northern and Western Europe, with a wide range of values. Eastern Europe has the lowest values, with a median of around 5 and lower protection rights for LGBT people.

```{r}
anova_result <- aov(LGBT..Policy.Index ~ region, data = merged_data)
summary(anova_result)
```

The ANOVA analysis shows a highly significant difference between the regions (p 2e-16), which means that the differences in the LGBT index between the regions are statistically significant.

Interpretation: LGBT policy differs systematically between the regions, with Western Europe and Northern Europe having significantly higher values than Eastern Europe and Southern Europe.

```{r}
cor_data <- merged_data %>%
  select(qc19, qc20, qc15_1, qc15_2, qc15_3,LGBT..Policy.Index) %>%
  mutate(across(everything(), as.numeric))  

cor_matrix <- cor(cor_data, use = "pairwise.complete.obs")

cor_matrix

ggcorrplot(cor_matrix, method = "circle", type = "lower",
           lab = TRUE, lab_size = 4, colors = c("red", "white", "blue"),
           title = "Correlation Matrix of Selected Variables")
             
```

The correlation table highlights interesting connections between societal views on LGBTI+ rights and the LGBT Policy Index. The strongest positive correlation appears between these variables: qc15_1 which gauges equal rights for LGB people; qc15_2 which measures acceptance of same-sex relationships; and qc15_3 which covers marriage equality. These values range from 0.75 to 0.82. It indicates that nations with high acceptance of LGB rights generally back marriage equality too. There's a moderate correlation between qc19 which covers support for transgender document changes and qc20 which deals with a third gender option, clocking in at 0.59. This suggests that countries with advanced laws for trans people often favor a third gender option on official documents as well. But the correlations between qc19/qc20 and general LGBTI+ attitudes ‒ qc15\_\* ‒ are moderate (0.33–0.43). A positive attitude towards LGB rights often occurs alongside supportive views on trans issues.

```{r}

lm_model <- lm(LGBT..Policy.Index ~ qc19 + qc20 + qc15_1 + qc15_2 + qc15_3, data = merged_data)

summary(lm_model)

```

The results show that approval of LGBT rights is strongly associated with a higher LGBT Policy Index. the strongest negative effects come from rejection of “marriage for all” (-1.76 points) and same-sex relationships (-0.94 points). neutral answers (“don't know”) are often as negative as rejection - uncertainty is therefore a strong signal for less legal protection of LGBT+ rights. the transgender-specific questions (qc19, qc20) are also significant, but with somewhat weaker effects than LGB rights. the model is significant, but only explains approx. 26.5% of the variance - other factors also play a role.

```{r}
# interaction model
lm_model_interact <- lm(LGBT..Policy.Index ~ qc19 * region + qc20 * region, data = merged_data)

summary(lm_model_interact)

```

In principle, the effects of qc19 and qc20 are relatively low in Eastern Europe, but uncertainty (“I don't know”) leads to stronger negative effects than rejection. regionNorthern Europe = +3.17\* Northern Europe has on average 3.17 points higher LGBT rights than Eastern Europe.\
regionSouthern Europe = +3.23\* Southern Europe has on average 3.23 points higher LGBT rights than Eastern Europe.\
regionWestern Europe = +4.19\* Western Europe has on average 4.19 points higher LGBT rights than Eastern Europe. Western Europe is the most progressive, followed by Southern Europe and Northern Europe, while Eastern Europe has the lowest index as a reference category.

In Northern & Southern Europe, disagreement with qc19 is strongly associated with lower LGBT rights - much more so than in Eastern Europe or Western Europe. In Western Europe, approval or disapproval of qc19 is not as important for the LGBT Policy Index, which means that other factors are more decisive there. In Northern Europe, rejection or uncertainty about a third option is associated with lower LGBT rights. In Southern Europe, uncertainty about qc20 has the opposite effect - there the discussion about the topic could be a sign of progress. In Western Europe, these issues hardly play a role in the LGBT index, as legal protection is high anyway. R² = 0.4341 43.4% of the variance in the LGBT Policy Index is explained by the model. p-value \< 2.2e-16 The model is highly significant.The new model explains 17% more variance, i.e. taking the region into account makes the model significantly better.






## Macro/micro

Macro level (country-specific factors)
Monarchy/Republic (from eu_gov_styles_table)
Bloc (from eu_blocs)
internac_trade (international trade)
trade_tariffs (trade tariffs)
trade_agreements (trade agreements)
online_purchases (online purchases)
AIC (Adjusted Individual Consumption)
unemployment_rate (unemployment rate)
min_wage (minimum wage)
LGBT..Policy.Index (Index for LGBT+ rights)
hdi (humand development index)
region 
country


Micro-level (individual factors)
d10 (gender)
D15b:   last occupation
d25: rural vs urban 
gen3, gen4, gen5, gen6 (generational affiliation)
d1r2 (self-assessment of political ideology)
d1r2 * d10 (Interaction: political ideology & gender)
sd3_recode (religiosity, recoded from s3)
d11 (age of the respondent)
d8 (level of education)
qc19 (Trans people should be allowed to change official documents)
qc20 (Third gender option in official documents)
qc15_1 (LGB people should have the same rights as heterosexuals)
qc15_2 (Same-sex relationships are normal)
qc15_3 (Same-sex marriage should be allowed in Europe)
qc4_7 (Perceived discrimination at work because of gender)
qc4_8 (Perceived job discrimination because of gender identity)
qc5_3 (Commitment in an organization against discrimination)
qc1_4 (Perceived discrimination due to sexual orientation)
qc1_8 (Perceived discrimination against trans people)
qc2_4 (Personal experience of discrimination based on sexual orientation)
qc2_6 (Personal experience of discrimination because of trans identity)
qc17_3 (Schools should provide information about sexual diversity)
qc17_4 (Schools should provide information about gender identity)
qc18_2 (Two men can show affection in public)
qc18_3 (Two women can show affection in public)


Data is from: https://hdr.undp.org/data-center/human-development-index#/indicies/HDI

```{r}
hdi_data <- read_excel("hdr-data.xlsx", sheet = 1)  

hdi_data <- hdi_data |> filter(year=="2019" & indicatorCode == "hdi")

merged_data

hdi_data <- hdi_data %>%
  rename(country_name = country)


merged_data <- merged_data %>%
  left_join(hdi_data, by = "country_name")

summary(merged_data)

```



```{r}
micro_data <- merged_data %>%
  select(id,
   qc20, qc15_1, qc15_2, qc15_3, qc4_7, qc4_8, qc5_3
  )


macro_data <- merged_data %>%
  select(id, country_name, LGBT..Policy.Index, region, value
  )

```




