---
title: "Irantzu_survey"
author: "Irantzu Lamarca"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  markdown: 
    wrap: sentence
---

# Societal attitudes toward LGBTI rights

## Libraries

```{r}
library(haven)
library(dplyr)
library(tidyverse)
library(readr)
library(ggplot2)
library(ggcorrplot)
library(nnet)
library(reshape2)
```

## Importing the dataset and preparing the data for analysis

```{r}
# Import the dataset 
data <- read_dta("ZA7575.dta")

# Extract variable labels
column_info <- data.frame(
  column_name = colnames(data),
  column_label = sapply(data, function(x) attr(x, "label"))
)

# Verify variable labels
print(column_info)

# Create country name variable
data <- data %>%
  mutate(country_name = case_when(
    q1_1  == 1 ~ "Belgium",
    q1_2  == 1 ~ "Denmark",
    q1_3  == 1 ~ "Germany",
    q1_4  == 1 ~ "Greece",
    q1_5  == 1 ~ "Spain",
    q1_6  == 1 ~ "France",
    q1_7  == 1 ~ "Ireland",
    q1_8  == 1 ~ "Italy",
    q1_9  == 1 ~ "Luxembourg",
    q1_10 == 1 ~ "Netherlands",
    q1_11 == 1 ~ "Portugal",
    q1_12 == 1 ~ "United Kingdom",
    q1_13 == 1 ~ "Austria",
    q1_14 == 1 ~ "Sweden",
    q1_15 == 1 ~ "Finland",
    q1_16 == 1 ~ "Cyprus",
    q1_17 == 1 ~ "Czech Republic",
    q1_18 == 1 ~ "Estonia",
    q1_19 == 1 ~ "Hungary",
    q1_20 == 1 ~ "Latvia",
    q1_21 == 1 ~ "Lithuania",
    q1_22 == 1 ~ "Malta",
    q1_23 == 1 ~ "Poland",
    q1_24 == 1 ~ "Slovakia",
    q1_25 == 1 ~ "Slovenia",
    q1_26 == 1 ~ "Bulgaria",
    q1_27 == 1 ~ "Romania",
    q1_28 == 1 ~ "Croatia",
    q1_29 == 1 ~ "Other countries",
    q1_30 == 1 ~ "DK",
    TRUE ~ NA_character_
  ))

```

## Checking for the useful variables for societal attitudes

Taking into account what has been previously analyzed, these are the variables necessary to carry out the descriptive analysis based on societal attitudes.

Key Variables Explaining Support for Transgender People (cq19) - cq19: Opinion on whether transgender people should be able to change official documents.
- country / country_name: Country code or name of the respondent's country (for country analysis).
- d11: Respondent's exact age (key socio-demographic factor).
- d10: Gender of the respondent (Male/Female).
- d8: Level of education (impact of education on support for trans rights).
- qc1_4: Perceived discrimination based on sexual orientation in the country.
- qc1_8: Perceived discrimination against transgender people in the country.
- qc2_4: Personal experience of discrimination based on sexual orientation.
- qc2_6: Personal experience of discrimination for being trans.
- qc4_8: Perceived disadvantage in employment due to gender identity.

Variables Measuring Social Attitudes towards LGBTI+ Rights - qc15_1: Opinion on whether LGB people should have the same rights as heterosexuals.
- qc15_2: Opinion on whether same-sex relationships are normal.
- qc15_3: Opinion on whether same-sex marriage should be allowed in Europe.
- qc17_3: Opinion on whether schools should inform about diversity in sexual orientation.
- qc17_4: Opinion on whether schools should report on diversity in gender identity.
- qc18_2: Opinion on whether two men can show affection in public without a problem.
- qc18_3: Opinion on whether two women can show affection in public without a problem.
- qc20: Opinion on whether official documents should include a third gender option.

```{r}
# Select key variables for analysis
data_clean <- data %>%
  select(country_name,
         qc19,  # Support for trans people in official documents
         d11,   # Age
         d10,   # Gender
         d8,    # Educational level 
         qc1_4, qc1_8, qc2_4, qc2_6, qc4_8,  # Discrimination and perception of inequality
         qc15_1, qc15_2, qc15_3,  # Opinions on LGB rights
         qc17_3, qc17_4, qc18_2, qc18_3,  # Diversity attitudes
         qc20)  # Opinion on third gender option in official documents

#Add ID variable 
data_clean <- data_clean %>%
  mutate(id = row_number()) %>%
  relocate(id, .before = qc19)  

head(data_clean)

# Transform variables into factors with descriptive labels
data_clean <- data_clean %>%
  mutate(
    qc19 = factor(qc19, levels = c(1,2,3),
                  labels = c("Yes", "No", "DK")),
    d10 = factor(d10, levels = c(1,2),
                 labels = c("Man", "Woman")),
    d8 = factor(d8),  # keep education as a factor 
    # 2 level variables
    qc2_4 = factor(qc2_4, levels = c(0,1), labels = c("No", "Yes")),
    qc2_6 = factor(qc2_6, levels = c(0,1), labels = c("No", "Yes")),
    qc4_8 = factor(qc4_8, levels = c(0,1), labels = c("No", "Yes")),
    # 3 level variables
    qc20 = factor(qc20, levels = c(1,2,3), labels = c("Yes", "No", "DK")),
    # 5 level variables
    qc15_1 = factor(qc15_1, levels = c(1,2,3,4,5),
                    labels = c("Totally agree", "Tend to agree", "Tend to disagree", "Totally disagree", "DK")),
    qc15_2 = factor(qc15_2, levels = c(1,2,3,4,5),
                    labels = c("Totally agree", "Tend to agree", "Tend to disagree", "Totally disagree", "DK")),
    qc15_3 = factor(qc15_3, levels = c(1,2,3,4,5),
                    labels = c("Totally agree", "Tend to agree", "Tend to disagree", "Totally disagree", "DK")),
    qc17_3 = factor(qc17_3, levels = c(1,2,3,4,5),
                    labels = c("Totally agree", "Tend to agree", "Tend to disagree", "Totally disagree", "DK")),
    qc17_4 = factor(qc17_4, levels = c(1,2,3,4,5),
                    labels = c("Totally agree", "Tend to agree", "Tend to disagree", "Totally disagree", "DK")),
    # 6 level variables 
    qc1_4 = factor(qc1_4, levels = c(1,2,3,4,5,6),
                   labels = c("Very widespread", "Fairly widespread", "Fairly rare", "Very rare", "Non-existent", "DK")),
    qc1_8 = factor(qc1_8, levels = c(1,2,3,4,5,6),
                   labels = c("Very widespread", "Fairly widespread", "Fairly rare", "Very rare", "Non-existent", "DK")),
    # 12 level variables
    qc18_2 = factor(qc18_2, levels = c(1,2,3,4,5,6,7,8,9,10,11,12)),
    qc18_3 = factor(qc18_3, levels = c(1,2,3,4,5,6,7,8,9,10,11,12))
  )

# Check for NA's
sum(is.na(data_clean)) #no NA's 

summary(data_clean)

```

## Descriptive, inferential and predictive analysis

### General statistics from the dataset (all, general)

```{r}

n_respondents <- nrow(data_clean) # Total number of observations
n_countries <- length(unique(data_clean$country_name)) # Number of countries 

cat("Total observations", n_respondents, "\n")
cat("Total countries", n_countries, "\n")

# Summary of key numerical variables
summary(data_clean[, c("d11")])  # Age

```

The data set analyzed includes a total of **27,438 respondents** from **28 countries**, which provides a representative sample to study attitudes towards supporting trans people in obtaining official documents.Regarding the variable **age (`d11`)**, a wide distribution is observed, with an age range between **15 and 98 years**.
The mean age of the respondents is **51.56 years**, indicating that the sample is mostly composed of older adults.
In addition, the median age is **53 years**, suggesting that half of the respondents are **53 years or younger**, while the other half are older.
The **first quartile (37 years old)** and **third quartile (66 years old)** show that 50% of the sample is in an age range between **37 and 66 years old**, which allows us to analyze generational differences in support for trans rights.
The age of respondents is a key factor in this analysis, as attitudes towards LGBTI rights can vary significantly by age group, with younger generations generally showing greater acceptance in previous studies.

### Distribution of the support for trans people (qc19) (all, general)

```{r}

table_qc19 <- data_clean %>%
  count(qc19) %>%
  mutate(percentage = n / sum(n) * 100)

print(table_qc19)

ggplot(data_clean, aes(x = qc19, fill = qc19)) +
  geom_bar() +
  labs(title = "Transgender Support Distribution",
       x = "Support level", y = "Number of answers") +
  theme_minimal() +
  scale_fill_brewer(palette = "Blues")


#this graph could be added to the part of Sophie, the one that says "Transgender Rights in Europe: Which Countries Support Legal Gender Changes?"
```

The survey results regarding support for transgender individuals obtaining official documents (`qc19`) reveal key insights into public opinion.
A **majority (52.71%)** of respondents express support for this right, indicating a general positive attitude toward transgender inclusivity in legal documentation.
However, **35.33% oppose** the idea, showing that a significant portion of the population still holds reservations or negative views on the issue.
Additionally, **11.95% of respondents answered "Don't Know (DK)"**, suggesting a level of uncertainty or lack of awareness regarding the topic.

These results highlight a **clear, but not overwhelming, majority in favor of transgender rights**, which may reflect broader trends in social acceptance and legal progress in different countries.
However, the **considerable opposition (over one-third of respondents)** suggests that cultural, political, and religious factors likely play a role in shaping attitudes toward transgender individuals.
The presence of nearly **12% undecided respondents** indicates that educational initiatives and policy discussions could influence public opinion further.

### Differences in support by sociodemographic characteristics

```{r}
table_qc19_gender <- data_clean %>%
  group_by(d10, qc19) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100)

print(table_qc19_gender)

ggplot(data_clean, aes(x = qc19, fill = d10)) +
  geom_bar(position = "dodge") +
  labs(title = "Support for Transgender People by Gender",
       x = "Support level", y = "Number of responses", fill = "Gender") +
  theme_minimal()

```

The data reveals a clear gender-based difference in support for transgender individuals obtaining official documents.
On the one hand, 54.64% of women support trans rights, compared to 50.40% of men.
Higher support among women may be linked to greater empathy and openness to gender issues, which has been observed in previous research on LGBTQ+ rights.
On the other hand, men have a higher opposition rate (38.68%) compared to women (32.53%), indicating that men are more likely to reject transgender rights.
Men's higher opposition rates could be influenced by traditional gender norms and conservative social attitudes that are more prevalent among male respondents.
Adittionally, the percentage of respondents who are uncertain is slightly higher among women (12.81%) than among men (10.91%).The uncertainty among women suggests that some female respondents may be undecided rather than strictly opposed, indicating a potential opportunity for awareness and education efforts.
In conclusion, the results highlight a gender gap in attitudes towards transgender rights, with women being more supportive and men showing more resistance.

We will perform a chi-square test to know if these gender differences are statistically significant or not.

```{r}
# Contingency table
data_matrix <- matrix(c(6296, 8167,  # Yes responses
                        4832, 4863,  # No responses
                        1364, 1916), # DK responses
                      nrow = 2, byrow = TRUE)

rownames(data_matrix) <- c("Man", "Woman")
colnames(data_matrix) <- c("Yes", "No", "DK")

#Chi-square
chi_square_test <- chisq.test(data_matrix)

print(chi_square_test)

```

Since p-value \< 0.05, we reject the null hypothesis, which means that there is a statistically significant relationship between gender and support for transgender rights.
The large X-squared value (2117) suggests that gender plays a strong role in shaping attitudes towards transgender individuals obtaining official documents.
This confirms the pattern observed in the descriptive analysis: women are significantly more likely to support transgender rights than men, while men show a higher rate of opposition.
So, we could say that the differences in support for transgender rights between men and women are not random, but statistically significant and that women are significantly more supportive, while men show more opposition.

### Exploration of perceived discrimination

```{r}
table_discrimination <- data_clean %>%
  count(qc1_4, qc1_8) %>%
  mutate(percentage = n / sum(n) * 100)

print(table_discrimination)

ggplot(data_clean, aes(x = qc1_4, fill = qc1_8)) +
  geom_bar(position = "fill") +
  labs(title = "Perceived Discrimination: Sexual Orientation vs. being Trans",
       x = "Perception of Discrimination based on Sexual Orientation",
       y = "Proportion", fill = "Discrimination against trans people") +
  theme_minimal()

```

The provided data examine the relationship between perceived discrimination based on sexual orientation (qc1_4) and being transgender (qc1_8).
The key findings reveal significant patterns of perceived discrimination across different categories.
There is higher perceived discrimination for sexual orientation and transgender identity, since a significant proportion of respondents perceive discrimination against both LGB individuals and transgender people as "Very Widespread" or "Fairly Widespread".The most frequent response combination is "Fairly Widespread" for both categories, with 19% of respondents selecting this.
Moreover, 7.96% of respondents believe that discrimination against both groups is "Very Widespread", reinforcing the idea that prejudice remains a serious issue in many societies.
There is also a perception of "Non-Existent" Discrimination as only a small fraction of respondents believe that discrimination against either LGB or transgender individuals does not exist, exactly 1.58% of respondents consider that discrimination against trans people is non-existent, whereas 0.98% say the same for sexual orientation.
This suggests that while some individuals do not acknowledge discrimination, they are a minority in the dataset.In addition, a notable proportion of respondents are uncertain about the level of discrimination.The highest rate of "Don't Know" responses appears when respondents also selected "Don't Know" for sexual orientation (7.03%), suggesting that awareness of discrimination issues may still be limited for some people.

We will continue by identifying countries with the highest perceived discrimination and exploring correlations between discrimination perceptions and demographic variables.

```{r}
#Countries with highest perceived discrimination
# Filter data: Perceived discrimination as "Very Widespread" or "Fairly Widespread"
data_discrimination <- data_clean %>%
  filter(qc1_4 %in% c("Very widespread", "Fairly widespread") | 
         qc1_8 %in% c("Very widespread", "Fairly widespread")) %>%
  group_by(country_name) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) 

# Normalize by total respondents per country
total_per_country <- data_clean %>%
  group_by(country_name) %>%
  summarise(total = n())

# Merge the counts and calculate the percentage
data_discrimination <- merge(data_discrimination, total_per_country, by = "country_name") %>%
  mutate(percentage = (count / total) * 100)


ggplot(data_discrimination, aes(x = reorder(country_name, -percentage), y = percentage)) +
  geom_bar(stat = "identity", fill = "red") +
  coord_flip() +
  labs(title = "Countries with Highest Perceived Discrimination",
       x = "Country", y = "Percentage of Respondents") +
  theme_minimal()


print(data_discrimination)

```

The analysis of perceived discrimination across countries reveals significant disparities in public opinion.Based on the data, Greece (76.29%) and Italy (75.16%) report the highest levels of perceived discrimination, followed closely by France (74.92%) and Portugal (69.54%).
This suggests that in these countries, a significant proportion of respondents believe discrimination against LGBTI individuals is widespread.
Conversely, countries such as Slovakia (23.98%), Estonia (31.34%), Latvia (32.25%), and Bulgaria (32.39%) report the lowest levels of perceived discrimination.
This could indicate either lower actual discrimination levels or differing social perceptions and awareness regarding discrimination issues.
The bar chart visually reinforces these findings, with Southern European countries tending to report higher discrimination levels, while Eastern European and Baltic nations report lower levels.

###Clustering analysis

```{r}
# Matrix with country variables
data_numeric <- data_clean %>%
  group_by(country_name) %>%
  summarise(support_trans = mean(as.numeric(qc19), na.rm = TRUE),
            discrimination_LGBTI = mean(as.numeric(qc1_4), na.rm = TRUE),
            discrimination_trans = mean(as.numeric(qc1_8), na.rm = TRUE))

# Clustering
set.seed(123)
kmeans_result <- kmeans(scale(data_numeric[, -1]), centers = 3)

data_numeric$cluster <- factor(kmeans_result$cluster)

ggplot(data_numeric, aes(x = reorder(country_name, support_trans), y = support_trans, fill = cluster)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(title = "Clustering of countries according to support for transgender rights",
       x = "Country", y = "Support level") +
  theme_minimal()

```

The clustering analysis of countries based on support for transgender rights reveals three distinct groups.
The **first cluster (red)** consists of countries such as **Hungary, Bulgaria, Romania, and Poland**, which show the lowest levels of support.
These nations are known for their more conservative social attitudes and have faced criticism for policies restricting LGBTQ+ rights.
The **second cluster (green)** includes countries like **Luxembourg, Austria, Estonia, Cyprus, and Slovakia**, which exhibit moderate levels of support, suggesting a transitional stance where societal attitudes might be shifting toward greater acceptance.
Finally, the **third cluster (blue)** comprises countries with the highest support for transgender rights, including **Spain, the Netherlands, Denmark, France, and Sweden**, which are widely recognized for their progressive policies and strong legal protections for LGBTQ+ individuals.
This clustering highlights a clear **East-West divide**, where **Western and Northern European countries tend to be more supportive**, while **Central and Eastern European nations exhibit more resistance to transgender rights**.
The findings align with broader trends in societal attitudes towards LGBTQ+ rights, influenced by political, religious, and cultural factors across different regions.

###Correlations

We will compute now the correlations between discrimination perceptions and demographic variables.

```{r}
# Convert categorical variables to numeric 
data_numeric <- data_clean %>%
  mutate(
    qc1_4_num = as.numeric(factor(qc1_4, levels = c("Very widespread", "Fairly widespread", "Fairly rare", "Very rare", "Non-existent", "DK"))),
    qc1_8_num = as.numeric(factor(qc1_8, levels = c("Very widespread", "Fairly widespread", "Fairly rare", "Very rare", "Non-existent", "DK"))),
    d11 = as.numeric(d11),
    d10_num = as.numeric(factor(d10, levels = c("Man", "Woman"))),  
    d8_num = as.numeric(d8)  
  )

# Correlation matrix
cor_matrix <- cor(data_numeric[, c("qc1_4_num", "qc1_8_num", "d11", "d10_num", "d8_num")], use = "complete.obs")

print(cor_matrix)

ggcorrplot(cor_matrix, lab = TRUE, colors = c("blue", "white", "red"))

```

The correlation analysis between perceived discrimination and demographic variables provides insightful findings regarding their relationships.They reveal a strong positive correlation (0.64) between perceived discrimination based on sexual orientation and perceived discrimination towards transgender individuals.
This suggests that individuals who perceive high levels of discrimination against one group also tend to perceive high levels of discrimination against the other, indicating a shared perception of marginalization across these identities.
The correlation between age (**d11**) and perceived discrimination is positive but weak (0.19 with **qc1_4_num** and 0.20 with **qc1_8_num**).
This suggests that older individuals tend to report slightly higher levels of perceived discrimination compared to younger individuals, but the effect is not strong.
There is a moderate negative correlation between education level (**d8_num**) and perceived discrimination (-0.38 with **d11**, -0.08 with **qc1_8_num**, and -0.07 with **qc1_4_num**).
This indicates that individuals with higher levels of education tend to report lower perceptions of discrimination, possibly reflecting a greater awareness of anti-discrimination policies or personal social positioning that mitigates exposure to discrimination.
The correlation between gender (**d10_num**) and discrimination perceptions is near zero, suggesting that gender does not significantly influence how individuals perceive discrimination against LGBTQ+ groups in this dataset.
The results highlight that perceptions of discrimination towards different LGBTQ+ groups are interconnected, with those who recognize discrimination against one group likely acknowledging it for another.
Older individuals report slightly higher discrimination perceptions, while higher educational attainment appears to be associated with lower perceptions of discrimination.
Gender does not seem to play a crucial role in shaping discrimination perceptions.

We will continue with some regression models to predict support for transgender individuals obtaining official documents (qc19) based on various demographic and attitudinal factors.

### Regression

First, since qc19 is a categorical variable with three levels (Yes, No, Don't Know), we will use multinomial logistic regression.

```{r}

data_clean$qc19 <- factor(data_clean$qc19, levels = c("Yes", "No", "DK")) # Convert qc19 into a factor
# Run multinomial logistic regression model
model_multinom <- multinom(qc19 ~ d10 + d11 + d8 + qc1_4 + qc1_8, data = data_clean)

summary(model_multinom)

exp(coef(model_multinom))

```

The main results of the regression are as follows:

-   Gender Differences:The coefficient for d10Woman suggests that being a woman is associated with a higher likelihood of supporting transgender rights compared to men. This aligns with previous literature indicating that women tend to be more supportive of LGBTQ+ rights.
-   Age Influence (d11): The coefficient for d11 suggests that an increase in age slightly reduces the probability of supporting transgender rights. This indicates that older individuals may be more resistant to policy changes regarding gender inclusivity.
-   Education Level (d8 & d82 to d89): Higher education levels appear to have a significant positive impact on support for transgender rights. Respondents with more education are more likely to express support. Conversely, lower education levels are associated with a higher probability of being against or uncertain about transgender-inclusive policies.
-   Perception of Discrimination (qc1_4, qc1_8): Respondents who perceive discrimination based on sexual orientation or gender identity as widespread are more likely to support transgender rights. Those who believe discrimination is rare or non-existent tend to oppose transgender-inclusive policies.
-   Odds Ratios (Exp(Coef)): The exponentiated coefficients show how much more (or less) likely a respondent is to choose one category over another. Some of the strongest predictors include perceptions of discrimination (qc1_4, qc1_8) and education level (d8).

Now, we will finish our analysis by performing some correlations between some last key variables.

```{r}
data_numeric <- data_clean %>%
  mutate(across(where(is.factor), as.numeric))

# Correlation matrix
cor_matrix <- cor(data_numeric[, c("qc19", "qc1_4", "qc1_8", "qc15_1", "qc15_2", "qc15_3")], use = "pairwise.complete.obs")

print(cor_matrix)

melted_cor <- melt(cor_matrix)

ggplot(melted_cor, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile() +
  scale_fill_gradient2(low="red", high="blue", mid="white", midpoint=0) +
  theme_minimal() +
  labs(title="Map of Correlations between Key Variables")

```

The correlation analysis among key variables provides valuable insights into the relationships between different perceptions and opinions.
The strongest correlations are observed between qc15_1 (LGB rights statement: same rights as heterosexuals) and qc15_2 (LGB rights statement: nothing wrong with same-sex relationships), with a coefficient of 0.765, as well as between qc15_2 and qc15_3 (LGB rights statement: allow same-sex marriage throughout Europe) at 0.817.
These high correlations indicate that individuals who support equal rights for LGB individuals also tend to support same-sex marriage and affirm that there is nothing wrong with same-sex relationships.
Furthermore, qc1_4 (perceived discrimination based on sexual orientation) and qc1_8 (perceived discrimination based on being transgender) exhibit a moderate correlation of 0.638, suggesting that those who perceive widespread discrimination against individuals based on their sexual orientation also tend to perceive discrimination against transgender individuals.
This alignment indicates that public perceptions of discrimination against different marginalized groups are interrelated.
The correlations between qc19 (support for a third gender option in official documents) and other variables are weaker but still notable, particularly with qc15_1 (0.408) and qc15_2 (0.414), implying that individuals who favor LGB rights also tend to support the inclusion of a third gender option.
However, the correlations are not extremely strong, indicating that while these attitudes are related, they are not entirely dependent on one another.
Overall, the analysis suggests that perceptions of discrimination and support for LGBTQ+ rights are interconnected, but with varying degrees of association.
The strongest relationships are observed between different aspects of LGB rights, while perceived discrimination and support for transgender rights also show moderate connections.
These findings highlight the importance of considering intersectional perspectives when analyzing public attitudes toward discrimination and equality.
