---
title: "Group Presentation"
subtitle: "Survey Methodology II"
date: 2025-03-19
author: "Irantzu, Sophie, Diego, Gür"
format: 
  revealjs:
    theme: simple
    transition: slide
    transition-speed: fast
    embed-resources: TRUE
    slide-number: c/t
editor: visual
logo: https://www.rstudio.com/wp-content/uploads/2018/10/RStudio-Logo-Flat.png
---

```{r}
knitr::opts_chunk$set(error = TRUE, warning = TRUE, message = TRUE)
```

## Data cleaning

-   Libraries and cleaning Data

-   Creating a data frame with column names and labels

-   Convert into tidy Data

```{r Data cleaning, message=FALSE, warning=FALSE}
#| echo: false
library(haven)
library(dplyr)
library(ggplot2)
library(stringr) 
library(lubridate)
library(corrr)
library(ggcorrplot)
library(tidyr)
library(scales)
library(mice)
library(readxl)
library(leaflet)
library(sf)
library(RColorBrewer)
library(rnaturalearth)
library(rnaturalearthdata)
library(corrplot)
library(lme4) 
library(car)
library(nnet)
library(reshape2)
library(pROC)

data <- read_dta("ZA7575.dta")
column_info <- data.frame(
  column_name = colnames(data),
  column_label = sapply(data, function(x) attr(x, "label"))
)

data <- data %>%
  mutate(country_name = case_when(
    q1_1  == 1 ~ "Belgium",
    q1_2  == 1 ~ "Denmark",
    q1_3  == 1 ~ "Germany",
    q1_4  == 1 ~ "Greece",
    q1_5  == 1 ~ "Spain",
    q1_6  == 1 ~ "France",
    q1_7  == 1 ~ "Ireland",
    q1_8  == 1 ~ "Italy",
    q1_9  == 1 ~ "Luxembourg",
    q1_10 == 1 ~ "Netherlands",
    q1_11 == 1 ~ "Portugal",
    q1_12 == 1 ~ "United Kingdom",
    q1_13 == 1 ~ "Austria",
    q1_14 == 1 ~ "Sweden",
    q1_15 == 1 ~ "Finland",
    q1_16 == 1 ~ "Cyprus",
    q1_17 == 1 ~ "Czech Republic",
    q1_18 == 1 ~ "Estonia",
    q1_19 == 1 ~ "Hungary",
    q1_20 == 1 ~ "Latvia",
    q1_21 == 1 ~ "Lithuania",
    q1_22 == 1 ~ "Malta",
    q1_23 == 1 ~ "Poland",
    q1_24 == 1 ~ "Slovakia",
    q1_25 == 1 ~ "Slovenia",
    q1_26 == 1 ~ "Bulgaria",
    q1_27 == 1 ~ "Romania",
    q1_28 == 1 ~ "Croatia",
    q1_29 == 1 ~ "Other countries",
    q1_30 == 1 ~ "DK",
    TRUE ~ NA_character_  
  ))

data$country_name <- gsub("Czech Republic", "Czechia", data$country_name)
```


```{r echo=TRUE, results='hide'}
data <- data |> 
  mutate(
    qc19 = factor(qc19, levels = c(1, 2, 3), labels = c("Yes", "No", "DK")),
    d10 = case_when(
      d10 == 1 ~ 1,  # Man for best practice
      d10 == 2 ~ 0   # Woman
    ),
    d10 = factor(d10, levels = c(0, 1),
                 labels = c("Woman", "Man"))
  )

```

## Descriptive Analysis 

-   General descriptive analysis 

-   Legal framework

-   Historical background

-   Cultural and Societal norms

-   Economical analysis

## General descriptive analysis

-   Check for NA's 

-   Total observations 27438 

-   Total countries 28 

-   Target variable --> qc19: Opinion on whether transgender people should be able to change official documents

```{r}
#| echo: false
table_qc19 <- data |> 
  count(qc19) |> 
  mutate(percentage = n / sum(n) * 100)
```

##

```{r}

ggplot(data, aes(x = qc19, fill = qc19)) +
  geom_bar() +
  labs(title = "Transgender Support Distribution",
       x = "Support level", y = "Number of answers") +
  theme_minimal() +
  scale_fill_brewer(palette = "Blues")

```

## Which Countries Support Legal Gender Changes?

```{r}
#| echo: false

data <- data |> 
  mutate(qc19_numeric = case_when(
    qc19 == "Yes" ~ 1,  # Yes
    qc19 == "No" ~ 0,  # No
    qc19 == "DK" ~ NA_real_  # DK
  ))

selected_data_cultural <- data |> 
  mutate(qc19_numeric = case_when(
    qc19 == "Yes" ~ 1,  # Yes
    qc19 == "No" ~ 0,  # No
    qc19 == "DK" ~ NA_real_  # DK
  )) |> 
  select(
    qc19_numeric,
    sd3,
    gen3,
    gen4,
    gen5,
    gen6,
    d10,
    qc2t_3,
    d1r2
  )

qc19_summary_sorted <- data %>%
  group_by(country_name) %>%
  mutate(total_responses = n()) %>%  
  group_by(country_name, qc19_numeric) %>%
  summarise(count = n(), total_responses = first(total_responses), .groups = "drop") %>%
  mutate(percentage = (count / total_responses) * 100) %>%
  filter(qc19_numeric == "1") %>%
  arrange(percentage)
```


```{r}
ggplot(qc19_summary_sorted, aes(x = reorder(country_name, percentage), y = percentage, fill = percentage)) +
  geom_bar(stat = "identity", color = "black", width = 0.7) +
  coord_flip() +
  scale_fill_gradientn(
    colors = c("red", "orange", "yellow", "green", "blue", "purple"), 
    values = rescale(c(min(qc19_summary_sorted$percentage), max(qc19_summary_sorted$percentage))),
    guide = "colorbar"
  ) +
  geom_text(aes(label = paste0(round(percentage, 1), "%")), hjust = -0.2, size = 5, fontface = "bold") +
  labs(
    title = "Support for Transgender Document Change (qc19) per Country",
    x = "Country",
    y = "Support Rate (%)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 16),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    legend.position = "right"
  )

```

## Legal Framework

-   included the lgtb-rights index-data

-   index captures to which extent LGBT+ people have the same rights as straight and cisgender people

-   combines 18 individual policies, such as the legality of same-sex relationships, marriage, and gender marker changes

```{r}
#| echo: false
lgbt_index <- read.csv("lgbt-rights-index.csv")

lgbt_index <- lgbt_index |> filter(Year== 2019)

data <- data %>%
  left_join(lgbt_index, by = c("country_name" = "Entity"))
```

## LGBT Index by Country

```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")
  
map_data <- world %>%
  left_join(data, by = c("name" = "country_name"))

map_data$LGBT..Policy.Index <- as.numeric(as.character(map_data$LGBT..Policy.Index))

map_data <- map_data %>%
  filter(!is.na(LGBT..Policy.Index)) %>%
  group_by(admin) %>%
  slice(1) %>%  
  ungroup()

pal <- colorBin(
  palette = c("red", "orange", "yellow", "green", "blue", "purple"),
  domain = map_data$LGBT..Policy.Index,
  bins = 6,
  na.color = "transparent"
)

pmap <- leaflet(data = map_data) %>%
  addTiles() %>%
  addPolygons(
    fillColor = ~pal(LGBT..Policy.Index),
    color = "white",
    weight = 0.5,
    opacity = 1,
    fillOpacity = 0.7,
    highlight = highlightOptions(weight = 2, color = "black", bringToFront = TRUE),
    label = ~paste0(name, ": ", round(LGBT..Policy.Index, 1)),  # Entfernt das "%"
    labelOptions = labelOptions(direction = "auto")
  ) %>%
  addLegend(
    pal = pal,
    values = map_data$LGBT..Policy.Index,
    position = "topright",
    title = "LGBT Index",
    opacity = 1
  )

pmap
```

## Dividing Europe in West/North/South and East

```{r}
data <- data %>%
  mutate(region = case_when(
    country_name %in% c("Denmark", "Sweden", "Finland", "Estonia", "Latvia", "Lithuania") ~ "Northern Europe",
    country_name %in% c("Germany", "France", "Netherlands", "Belgium", "Luxembourg", 
                        "United Kingdom", "United Kingdom of Great Britain and Northern Ireland",  
                        "Ireland", "Austria") ~ "Western Europe",
    country_name %in% c("Spain", "Italy", "Portugal", "Greece", "Malta", "Cyprus") ~ "Southern Europe",
    country_name %in% c("Poland", "Hungary", "Czechia", "Slovakia", "Romania", 
                        "Bulgaria", "Croatia", "Slovenia") ~ "Eastern Europe",
    TRUE ~ "Other"
  ))

data <- data %>%
  mutate(LGBT..Policy.Index = as.numeric(as.character(LGBT..Policy.Index)))

ggplot(data, aes(x = region, y = LGBT..Policy.Index, fill = region)) +
  geom_boxplot() +
  labs(title = "LGBT Policy Index by Region",
       x = "Region", y = "LGBT Policy Index") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

             
```

## LGTB Index and Region

```{r echo=TRUE, results='hide'}

lm_model_interact <- lm(LGBT..Policy.Index ~ qc19 * region + qc20 * region, data = data)

summary(lm_model_interact)

```

-   North & South Europe: disagreement with qc19 is strongly associated with lower LGBT rights

-   Western Europe: the approval or disapproval of qc19 is not as important for the LGBT Policy Index -\> other factors are more important there

## Historical blocs

```{r message= FALSE, warning= FALSE,include=FALSE}


# socialist bloc - now EU countries
socialist_bloc <- c("Bulgaria", "Croatia", "Czechia", "Estonia", "Hungary", 
                    "Latvia", "Lithuania", "Poland", "Romania", "Slovakia", "Slovenia")

# western bloc - now EU countries
western_bloc <- c("Austria", "Belgium", "Cyprus", "Denmark", "Finland", 
                  "France", "Greece", "Ireland", "Italy", "Luxembourg", 
                  "Malta", "Netherlands", "Portugal", "Spain", "Sweden", "United Kingdom") # adding the UK because it was a member at the time of survey


mixed_bloc <- c("Germany") # putting Germany into mixed for obvious reasons

countries <- c(socialist_bloc, western_bloc, mixed_bloc)
blocs <- c(rep("Socialist", length(socialist_bloc)), rep("Western", length(western_bloc)), rep("Mixed", length(mixed_bloc)))

eu_blocs <- data.frame(Country = countries, Bloc = blocs)
print(eu_blocs)

eu_blocs$Bloc_binary <- ifelse(eu_blocs$Bloc == "Socialist", 0, 1)

# to match the country_name column for joining tables later

eu_blocs <-
  eu_blocs |> 
  rename(country_name = Country)


save(eu_blocs, file = "eu_blocs.rda")

data <- data |> 
  left_join(eu_blocs, by = "country_name")

summary_df_pol <- as.data.frame(table(data$qc19, data$Bloc))
colnames(summary_df_pol) <- c("qc19", "Bloc", "Count")

# Use the existing qc19 values which are "Yes", "No", "DK"
# No need for levels = c("1", "2", "3") since qc19 is already labeled
summary_df_pol$qc19_label <- factor(summary_df_pol$qc19, levels = c("Yes", "No", "DK"))


```

```{r warning= FALSE}

ggplot(summary_df_pol, aes(x = Bloc, y = Count, fill = qc19_label)) +
  geom_bar(stat = "identity", color = "black", size = 0.5) +
  scale_fill_manual(values = c("Yes" = "#a1d99b", "No" = "#F44336", "DK" = "#B0BEC5")) +
  labs(title = "Transgender Document Change Support-Historical Bloc", x = "Bloc", y = "Count", fill = "Response") +
  theme_minimal(base_size = 14) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"),
        legend.title.position = "top",
        legend.position = "top",
        legend.key.size = unit(0.3, "cm"),
        panel.grid.major.x = element_blank())

```

## Generations

```{r message= FALSE, warning= FALSE,include=FALSE}

# before plotting, let's summarize each generation and create percentages for support/rejection levels

gen_summary <- rbind(
  data |> 
    filter(gen3 == 1) |> 
    group_by(qc19) |> 
    summarise(Count = n()) |> 
    mutate(Generation = "Before 1946", Percentage = Count / sum(Count) * 100),
  data |> 
    filter(gen4 == 1) |> 
    group_by(qc19) |> 
    summarise(Count = n()) |> 
    mutate(Generation = "1946-1964", Percentage = Count / sum(Count) * 100),
  data |> 
    filter(gen5 == 1) |> 
    group_by(qc19) |> 
    summarise(Count = n()) |> 
    mutate(Generation = "1965-1980", Percentage = Count / sum(Count) * 100),
  data |> 
    filter(gen6 == 1) |> 
    group_by(qc19) |> 
    summarise(Count = n()) |> 
    mutate(Generation = "After 1980", Percentage = Count / sum(Count) * 100)
) |> 
  mutate(Generation = factor(Generation, levels = c("Before 1946", "1946-1964", "1965-1980", "After 1980")))


```

```{r warning= FALSE}
# visualizing

ggplot(gen_summary, aes(x = qc19, y = Percentage, fill = Generation)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = sprintf("%.1f%%", Percentage)), 
            position = position_dodge(width = 0.9), vjust = -1, size = 3) +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "LGBT Documentation Rights - Generation", x = "", y = "Support percentage", fill = "Generation") +
  theme_minimal()

```

## Religion

```{r message= FALSE, warning= FALSE,include=FALSE}


# Recoding all Muslim categories since. They are a minority and summing them up could be better
summary_df_rel <- data |> 
  mutate(sd3_recode = case_when(
    sd3 %in% c(6, 7, 8) ~ 7,  #  sunni, shia, and other into one muslim level
    TRUE ~ sd3
  )) %>%
  group_by(sd3_recode, qc19) |> 
  summarise(Count = n()) |> 
  group_by(sd3_recode) |> 
  mutate(Percentage = Count / sum(Count) * 100) |> 
  ungroup() |> 
  mutate(sd3_recode = factor(sd3_recode, levels = c(1, 2, 3, 4, 5, 7, 9, 10, 11, 12, 13, 14, 15, 16),
                             labels = c("Catholic", "Orthodox Christian", "Protestant", "Other Christian", "Jewish", "Muslim", "Sikh", "Buddhist", "Hindu", "Atheist", "Non-believer/Agnostic", "Other", "Refusal", "DK")))



```

```{r warning= FALSE}
# plotting
ggplot(summary_df_rel, aes(x = sd3_recode, y = Percentage, fill = qc19)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("Yes" = "#a0d6b3", "No" = "#e6886f", "DK" = "#B0BEC5")) +
  labs(title = "QC19 Responses by Religion/Beliefs",
       x = "Religion/Beliefs", y = "Percentage", fill = "Response") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Political Ideology - self reports

```{r}

d1r2_props <- data %>%
  group_by(d1r2) %>%
  summarise(count = n(), .groups = "drop") %>%
  mutate(proportion = count / sum(count),
         d1r2_label = factor(d1r2, levels = c(1, 2, 3, 4, 5, 9),
                             labels = c("(1-2) Left", "(3-4)", "(5-6) Centre", 
                                        "(7-8)", "(9-10) Right", "DK/Refusal"))) %>%
  select(d1r2_label, count, proportion)
```

```{r}
ggplot(d1r2_props, aes(x = d1r2_label, y = proportion, fill = d1r2_label)) +
  geom_bar(stat = "identity") +
  labs(title = "Political Ideology Self-Reports",
       x = "Political Orientation",
       y = "Proportion") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set3") + # Optional: Use a color palette
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Societal attitudes

-   Correlations between discrimination perceptions and demographic variables

```{r}
#| echo: false
# Convert categorical variables to numeric 
data <- data |> 
  mutate(
    d10_num = as.numeric(d10),
    qc1_4_num = as.numeric(qc1_4),
    qc1_8_num = as.numeric(qc1_8),
    d11 = as.numeric(d11),
    d8_num = as.numeric(d8)  
  )

cor_matrix <- cor(data[, c("qc1_4_num", "qc1_8_num", "d11", "d10_num", "d8_num")], use = "complete.obs")

```

```{r}
ggcorrplot(cor_matrix, lab = TRUE, colors = c("blue", "white", "red"))
```


## Economic variables

-   Economic questions from the survey: internac_trade = qa1, trade_tariffs = qa12, trade_agreements = qa13, online_purchases = qa14, occupation = d15a, urban_rural = d25 or social_class = d63.

-   3 new external variables per country:

    /- Actual individual consumption (AIC): value of products actually consumed by individuals.

    /- Unemployment rate

    /- Minimum wage

```{r include=FALSE}

#| message: false
#| warning: false

# Let´s first select our variables of interest:
economic_data <- data |> select(
  "qc19", "country_name", "d11", "qa1", "qa12", "qa13", "qa14", "qa17","d15a", "d15b", "d25", "d60", "d63") |> 
  rename(age = d11, 
         internac_trade = qa1,
         trade_tariffs = qa12,
         trade_agreements = qa13,
         online_purchases = qa14,
         trust_EU_trade = qa17,
         occupation = d15a,
         past_occupation = d15b,
         urban_rural = d25,
         financial_difficulties = d60,
         social_class = d63
     )



# We will get the Actual Individual Consumption metric

# While GDP is mainly an indicator of the level of economic activity, actual individual consumption (AIC) is an alternative indicator better adapted to describe the material welfare of households. found here (https://ec.europa.eu/eurostat/statistics-explained/index.php?title=GDP_per_capita,_consumption_per_capita_and_price_level_indices#Data_sources)

AIC_per_capita <- read_excel("AIC_per_capita.xlsx")

AIC_per_capita <- AIC_per_capita |> rename("countries"="TIME")

AIC_per_capita
# We will get the unemployment rate

unemployment_rate <- read_excel("unemployment_rate.xlsx")
unemployment_rate <-unemployment_rate |>  pivot_longer(-countries, names_to = "years", values_to = "unemployment_rate") |> filter(years == 2019) |> select(-years)
unemployment_rate$unemployment_rate <- as.numeric(unemployment_rate$unemployment_rate)

unemployment_rate
# We will get the minimum wage
minimum_wage <- read_excel("EU_minimum_wage.xlsx")
minimum_wage <- minimum_wage |> pivot_longer(-countries, names_to = "years", values_to = "min_wage")|> filter(years == 2019) |> select(-years)
minimum_wage$min_wage <- as.numeric(minimum_wage$min_wage)
minimum_wage

# Merge the economic variables: 

AIC_and_unemployment <- full_join(AIC_per_capita, unemployment_rate, by = "countries")
economic_variables <- full_join(AIC_and_unemployment,minimum_wage, by = "countries")
economic_variables

# Finally, we merge the original survey data with the new economic variables
economic_data <- economic_data |> left_join(economic_variables, by = c("country_name" = "countries"))

economic_data# Let´s first select our variables of interest:
economic_data <- data |> select(
  "qc19", "country_name", "d11", "qa1", "qa12", "qa13", "qa14", "qa17","d15a", "d15b", "d25", "d60", "d63") |> 
  rename(age = d11, 
         internac_trade = qa1,
         trade_tariffs = qa12,
         trade_agreements = qa13,
         online_purchases = qa14,
         trust_EU_trade = qa17,
         occupation = d15a,
         past_occupation = d15b,
         urban_rural = d25,
         financial_difficulties = d60,
         social_class = d63
     )



# We will get the Actual Individual Consumption metric

# While GDP is mainly an indicator of the level of economic activity, actual individual consumption (AIC) is an alternative indicator better adapted to describe the material welfare of households. found here (https://ec.europa.eu/eurostat/statistics-explained/index.php?title=GDP_per_capita,_consumption_per_capita_and_price_level_indices#Data_sources)

AIC_per_capita <- read_excel("AIC_per_capita.xlsx")

AIC_per_capita <- AIC_per_capita |> rename("countries"="TIME")

AIC_per_capita
# We will get the unemployment rate

unemployment_rate <- read_excel("unemployment_rate.xlsx")
unemployment_rate <-unemployment_rate |>  pivot_longer(-countries, names_to = "years", values_to = "unemployment_rate") |> filter(years == 2019) |> select(-years)
unemployment_rate$unemployment_rate <- as.numeric(unemployment_rate$unemployment_rate)

unemployment_rate
# We will get the minimum wage
minimum_wage <- read_excel("EU_minimum_wage.xlsx")
minimum_wage <- minimum_wage |> pivot_longer(-countries, names_to = "years", values_to = "min_wage")|> filter(years == 2019) |> select(-years)
minimum_wage$min_wage <- as.numeric(minimum_wage$min_wage)
minimum_wage

# Merge the economic variables: 

AIC_and_unemployment <- full_join(AIC_per_capita, unemployment_rate, by = "countries")
economic_variables <- full_join(AIC_and_unemployment,minimum_wage, by = "countries")
economic_variables

# Finally, we merge the original survey data with the new economic variables
economic_data <- economic_data |> left_join(economic_variables, by = c("country_name" = "countries"))

economic_data





```

## Multiple Imputation with MICE

-   Na´s in the min_wage variable

-   Most similar distribution is the lasso regression

```{r include=FALSE}
# I will change those Na with 0. In the past_occupation case, having a 0 will #signify that there was no previous occupation.
set.seed(123)
colSums(is.na(economic_data))

cols_to_replace <- c("past_occupation")


economic_data <- economic_data |> mutate(across(all_of(cols_to_replace), ~ replace_na(., 0)))

# the rest of the variables with NA´s are due to a lack of information of those countries. They are the external economic variables

# Let's perform multiple imputation on missing values in minimum wage because it reduces bias, preserves variability, and maintains relationships between variables. Unlike complete-case analysis, which can introduce systematic bias and reduce statistical power, MI generates multiple plausible datasets, capturing uncertainty in missing values. This approach is particularly important for economic variables, which are often correlated and not missing completely at random. 

# let´s do it with MICE


variables <- c("original", "imputed_pmm", "imputed_cart", "imputed_lasso")
titles <- c("Original Data", "Predictive Mean Matching", "Classification & Regression Trees", "Lasso Regression")
colors_fill <- c("black", "#E69F00", "#56B4E9", "#009E73")


min_wage_mice_imputed <- data.frame(
  original = economic_variables$min_wage,
  imputed_pmm = complete(mice(economic_variables, m=5, method = "pmm", seed=123))$min_wage,
  imputed_cart = complete(mice(economic_variables, m=5, method = "cart", seed=123))$min_wage,
  imputed_lasso = complete(mice(economic_variables, m=5, method = "lasso.norm", seed=123))$min_wage
)

# --- the most similar distribution for min_wage is the lasso regression
min_wage_mice_imputed_long <- min_wage_mice_imputed |> 
  pivot_longer(all_of(variables), names_to = "method", values_to = "value")






# Now let´s integrate the imputed values into the original table:

economic_variables <- economic_variables |> 
  mutate(
    min_wage = ifelse(is.na(min_wage), min_wage_mice_imputed$imputed_lasso, min_wage)
  )
data <- data |> left_join(economic_variables, by = c("country_name" = "countries"))


```

```{r}
#| warning: false
min_wage_mice_imputed_long |> 
  mutate(title = factor(method, levels = variables, labels = titles)) |> 

ggplot(aes(x = value, fill = title)) +
  geom_histogram(binwidth = 1, color = "#808080", position = "identity") +
  facet_wrap(~title, scales = "free_y") +
  scale_fill_manual(values = colors_fill) +
  theme_classic() +
  theme(legend.position = "none")
a <- ggplot(data, aes(x = as.factor(qa1), fill = as.factor(qc19))) +
  geom_bar(position = "dodge") +
  labs(title = "Support for Trans Rights by International Trade Benefit",
       x = "International Trade Benefit", 
       y = "Count", 
       fill = "Support for Trans Rights (qc19)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Graph: International Trade Benefit.

International trade benefit -\> - 1: yes, benefitting a lot - 2: yes, benefitting somewhat - 3: No, not really benefitting 4: No, not benefitting at all 5: DK.

```{r}
 
a
```

## Feature engineering + Multicollinearity

```{r}
#| echo: false
data_selected <- data |> select(
  country_name, region, LGBT..Policy.Index, 
  qc19_numeric, qc20,    
  qc15_1, qc15_2, qc15_3,    
  qc17_3, qc17_4, qc18_2, qc18_3,  
  qc1_4, qc1_8, qc2_4, qc2_6, qc4_7, qc4_8, qc5_3,
  d11, d15a, d15b, d25, d60, d63,  
  qa1, qa12, qa13, qa14, qa17,   
  unemployment_rate, min_wage, AIC,
  sd3, d1r2, d10, gen3, gen6
)

data_selected$country_name <- as.factor(data_selected$country_name)
data_selected$region <- as.factor(data_selected$region)

factor_vars <- names(data_selected)[sapply(data_selected, is.factor)]

chi_results <- numeric(length(factor_vars))
names(chi_results) <- factor_vars

for (var in factor_vars) {
  test <- chisq.test(table(data_selected[[var]], data_selected$qc19_numeric))
  chi_results[var] <- test$p.value  
}

chi_results <- sort(chi_results)

```

```{r}
#| echo: false
cor_matrix <- cor(data_selected[, sapply(data_selected, is.numeric)], use = "pairwise.complete.obs")
```


```{r, fig.width=12, fig.height=10, out.width="100%"}
ggcorrplot(cor_matrix, lab = TRUE, lab_size = 3, tl.cex = 8, insig = "blank")  + theme_minimal()
```


## Model

-   Step: AIC=8607.78 qc19_numeric ~ country_name + qc20 + qc15_1 + qc15_2 + qc15_3 + qc17_3 + qc17_4 + qc18_2 + qc18_3 + qc1_4 + qc4_7 + qc4_8 + qc5_3 + d11 + d15a + d15b + d60 + qa1 + qa14 + d10
-   Scaled numerical values and removed non-significant variables
-   Checked for NA's

```{r message= FALSE, warning= FALSE,include=FALSE}
#| echo: false
data_selected <- data_selected %>%
  select(-gen6, -gen3)

best_model <- MASS::stepAIC(glm(qc19_numeric ~ ., data = data_selected, family = binomial), direction = "both")


data_selected <- data_selected |> 
  mutate(
    d11_scaled = scale(d11),
    d15a_scaled = scale(d15a),
    d15b_scaled = scale(d15b),
    d60_scaled = scale(d60)
  )


# Let´s check for na´s

colSums(is.na(data_selected))

data_selected$qc19_numeric[is.na(data_selected$qc19_numeric)] <- median(data_selected$qc19_numeric,na.rm=TRUE)
data_selected <- data_selected |> select(-d60)
data_selected <- data_selected |> select(-d15b_scaled)

data_selected %>%
  select(where(is.numeric)) %>%
  correlate() %>%
  focus(d15b)
```

## Imputation

-   Replace NA's in target variable with median
-   Apply MICE for the other variables

```{r message= FALSE, warning= FALSE,include=FALSE}
#| echo: false
vars_for_imputation <- data_selected %>%
  select(d15b, d15a, d11_scaled, d60_scaled, d1r2)
str(vars_for_imputation)

vars_for_imputation <- vars_for_imputation %>%
  mutate(
    d11_scaled = as.numeric(d11_scaled),
    d60_scaled = as.numeric(d60_scaled)
)

vars_for_imputation <- vars_for_imputation %>%
  mutate(across(where(is.labelled),as.numeric))
imputed_data <- mice(vars_for_imputation, method = "pmm", m = 5, maxit = 50, seed = 123)
```


```{r}
summary(imputed_data)

```

```{r message= FALSE, warning= FALSE,include=FALSE}
#| echo: false
data_selected$d15b <- complete(imputed_data)$d15b

na_count <- colSums(is.na(data_selected))


na_percentage <- (na_count / nrow(data_selected)) * 100

na_data <- data.frame(Variable = names(na_count), NA_Count = na_count, NA_Percentage = na_percentage)
na_data <- na_data[na_data$NA_Count > 0, ]
print(na_data)

```

## GLMER Model

```{r message= FALSE, warning= FALSE,include=FALSE}
#| echo: false
data_selected <- head(data_selected, 1000)
```


```{r echo=TRUE, results='hide', message= FALSE, warning=FALSE}
glmer_model_alt <- glmer(qc19_numeric ~ d10 + d11_scaled + I(d11_scaled^2) + qc20 + 
                        qc15_1 + qc15_2 + qc15_3 + qc17_3 + qc17_4 + 
                        qc1_4 + qc4_7 + d15b + qa1 + (1 | country_name), 
                      data = data_selected, family = binomial,
                      control = glmerControl(optimizer = "nloptwrap", 
                                             optCtrl = list(maxfun = 300000)))
```


```{r message= FALSE, warning= FALSE}
summary(glmer_model_alt)

```

## 

-   Multicollinearity revised with VIF --> <2 no problem
-   Model precision: 
```{r}
#| echo: false
vif_values <- car::vif(glmer_model_alt)

pred_probs_glmer <- predict(glmer_model_alt, newdata = data_selected, type = "response")

pred_classes_glmer <- ifelse(pred_probs_glmer > 0.5, 1, 0)

conf_matrix_glmer <- table(data_selected$qc19_numeric, pred_classes_glmer)

accuracy_glmer <- sum(diag(conf_matrix_glmer)) / sum(conf_matrix_glmer)
precision_glmer <- conf_matrix_glmer[2,2] / sum(conf_matrix_glmer[,2])
recall_glmer <- conf_matrix_glmer[2,2] / sum(conf_matrix_glmer[2,])
f1_score_glmer <- 2 * ((precision_glmer * recall_glmer) / (precision_glmer + recall_glmer))
```


```{r}
library(knitr)
metrics <- data.frame(
  Metric = c("Accuracy", "Precision", "Recall", "F1-Score"),
  Value = round(c(accuracy_glmer, precision_glmer, recall_glmer, f1_score_glmer), 4)
)
kable(metrics)
```

## ROC / AUC

```{r message= FALSE, warning= FALSE,include=FALSE}
#| echo: false
# ROC y AUC
roc_glmer <- roc(data_selected$qc19_numeric, pred_probs_glmer)
auc_glmer <- auc(roc_glmer)
```


```{r}
# compute ROC curve
plot(roc_glmer, col = "blue", main = paste("ROC Curve - GLMER | AUC:", round(auc_glmer, 3)))
```


```{r}
#| echo: false
# Find the best threshold
optimal_threshold <- coords(roc_glmer, "best", ret = "threshold")

# Apply the new threshold
pred_classes_optimal <- ifelse(pred_probs_glmer > optimal_threshold, 1, 0)

```

## Random Forest
```{r message= FALSE, warning= FALSE,include=FALSE}
#| echo: false
library(randomForest)
```


```{r echo=TRUE, results='hide', message= FALSE, warning= FALSE}
rf_model <- randomForest(qc19_numeric ~ ., data = data_selected, ntree = 500, importance = TRUE)
```


```{r message= FALSE, warning= FALSE,include=FALSE}
#| echo: false
# Prediction and evaluation
rf_pred <- predict(rf_model, newdata = data_selected)
conf_matrix_rf <- table(data_selected$qc19_numeric, rf_pred)
accuracy_rf <- sum(diag(conf_matrix_rf)) / sum(conf_matrix_rf)
```


```{r message= FALSE, warning= FALSE}
varImpPlot(rf_model)
```

## 

-   Training / testing (80-20)

```{r message= FALSE, warning= FALSE,include=FALSE}
#| echo: false
# training testing
set.seed(123)

library(caret)
in_train <- createDataPartition(y = data_selected$qc19_numeric, p = 0.8, list = FALSE)

in_train <- in_train[, 1]

training <- data_selected[in_train, ]
testing <- data_selected[-in_train, ]

#remove rows with NA's
training <- na.omit(training)
testing <- na.omit(testing)
f_predictions <- predict(rf_model, testing, type = "response")

# Convert to binary classes (0.5 threshold)
rf_pred_classes <- ifelse(rf_pred > 0.5, 1, 0)

# confusion matrix
rf_predictions <- predict(rf_model, testing, type = "class")  # Debería ser "class" en lugar de "response"

rf_pred_classes <- ifelse(rf_predictions == 1, 1, 0)


conf_matrix_rf <- table(testing$qc19_numeric, rf_pred_classes)
print(conf_matrix_rf)

# Accuracy, Precision, Recall y F1-Score
accuracy_rf <- sum(diag(conf_matrix_rf)) / sum(conf_matrix_rf)
precision_rf <- conf_matrix_rf[2,2] / sum(conf_matrix_rf[,2])
recall_rf <- conf_matrix_rf[2,2] / sum(conf_matrix_rf[2,])
f1_score_rf <- 2 * ((precision_rf * recall_rf) / (precision_rf + recall_rf))
```

```{r message= FALSE, warning= FALSE,include=FALSE}
#| echo: false
library(knitr)
metrics_comparison <- data.frame(
  Model = c("Random Forest", "Our Model"),
  Accuracy = round(c(accuracy_rf, accuracy_glmer), 4),
  Precision = round(c(precision_rf, precision_glmer), 4),
  Recall = round(c(recall_rf, recall_glmer), 4),
  `F1-Score` = round(c(f1_score_rf, f1_score_glmer), 4)
)

```

```{r}
kable(metrics_comparison)
```

## Additional model with 2 more variables

```{r echo=TRUE, results='hide', message= FALSE, warning=FALSE}
glmer_model_alt2.0 <- glmer(qc19_numeric ~ d10 + d11_scaled + I(d11_scaled^2) + qc20 + 
                        qc15_1 + qc15_2 + qc15_3 + qc17_3 + qc17_4 + sd3 + d1r2 +
                        qc1_4 + qc4_7 + d15b + qa1 + (1 | country_name), 
                      data = data_selected, family = binomial,
                      control = glmerControl(optimizer = "nloptwrap", 
                                             optCtrl = list(maxfun = 300000)))
```


```{r}
#| echo: false
pred_probs_glmer2.0 <- predict(glmer_model_alt2.0, newdata = data_selected, type = "response")

pred_classes_glmer2.0 <- ifelse(pred_probs_glmer2.0 > 0.5, 1, 0)

conf_matrix_glmer <- table(data_selected$qc19_numeric, pred_classes_glmer2.0)

accuracy_glmer <- sum(diag(conf_matrix_glmer)) / sum(conf_matrix_glmer)
precision_glmer <- conf_matrix_glmer[2,2] / sum(conf_matrix_glmer[,2])
recall_glmer <- conf_matrix_glmer[2,2] / sum(conf_matrix_glmer[2,])
f1_score_glmer <- 2 * ((precision_glmer * recall_glmer) / (precision_glmer + recall_glmer))
```

```{r}
metrics <- data.frame(
  Metric = c("Accuracy", "Precision", "Recall", "F1-Score"),
  Value = round(c(accuracy_glmer, precision_glmer, recall_glmer, f1_score_glmer), 4)
)
kable(metrics)
```


