---
title: "Group Presentation"
subtitle: "Survey Methodology II"
date: 2025-03-19
author: "Irantzu, Sophie, Diego, Gür"
format: 
  revealjs:
    theme: simple
    transition: slide
    transition-speed: fast
    embed-resources: TRUE
    slide-number: c/t
editor: visual
logo: https://www.rstudio.com/wp-content/uploads/2018/10/RStudio-Logo-Flat.png
---

## Data cleaning

-   Libraries and cleaning Data

-   Creating a data frame with column names and labels

-   Convert into tidy Data

```{r Data cleaning}
#| echo: false

library(haven)
library(dplyr)
library(ggplot2)
library(stringr) 
library(lubridate)
library(corrr)
library(ggcorrplot)
library(tidyr)
library(scales)
library(mice)
library(readxl)
library(leaflet)
library(sf)
library(RColorBrewer)
library(rnaturalearth)
library(rnaturalearthdata)
library(corrplot)
library(lme4) 
library(car)
library(nnet)
library(reshape2)
library(pROC)

data <- read_dta("ZA7575.dta")
column_info <- data.frame(
  column_name = colnames(data),
  column_label = sapply(data, function(x) attr(x, "label"))
)

data <- data %>%
  mutate(country_name = case_when(
    q1_1  == 1 ~ "Belgium",
    q1_2  == 1 ~ "Denmark",
    q1_3  == 1 ~ "Germany",
    q1_4  == 1 ~ "Greece",
    q1_5  == 1 ~ "Spain",
    q1_6  == 1 ~ "France",
    q1_7  == 1 ~ "Ireland",
    q1_8  == 1 ~ "Italy",
    q1_9  == 1 ~ "Luxembourg",
    q1_10 == 1 ~ "Netherlands",
    q1_11 == 1 ~ "Portugal",
    q1_12 == 1 ~ "United Kingdom",
    q1_13 == 1 ~ "Austria",
    q1_14 == 1 ~ "Sweden",
    q1_15 == 1 ~ "Finland",
    q1_16 == 1 ~ "Cyprus",
    q1_17 == 1 ~ "Czech Republic",
    q1_18 == 1 ~ "Estonia",
    q1_19 == 1 ~ "Hungary",
    q1_20 == 1 ~ "Latvia",
    q1_21 == 1 ~ "Lithuania",
    q1_22 == 1 ~ "Malta",
    q1_23 == 1 ~ "Poland",
    q1_24 == 1 ~ "Slovakia",
    q1_25 == 1 ~ "Slovenia",
    q1_26 == 1 ~ "Bulgaria",
    q1_27 == 1 ~ "Romania",
    q1_28 == 1 ~ "Croatia",
    q1_29 == 1 ~ "Other countries",
    q1_30 == 1 ~ "DK",
    TRUE ~ NA_character_  
  ))

data$country_name <- gsub("Czech Republic", "Czechia", data$country_name)

data <- data |> 
  mutate(
    qc19 = factor(qc19, levels = c(1, 2, 3), labels = c("Yes", "No", "DK")),
    d10 = case_when(
      d10 == 1 ~ 1,  # Man for best practice
      d10 == 2 ~ 0   # Woman
    ),
    d10 = factor(d10, levels = c(0, 1),
                 labels = c("Woman", "Man"))
  )

```

## legal Framework

-   included the lgtb-rights index-data

-   index captures to which extent LGBT+ people have the same rights as straight and cisgender people

-   combines 18 individual policies, such as the legality of same-sex relationships, marriage, and gender marker changes

```{r}
#| echo: false
lgbt_index <- read.csv("lgbt-rights-index.csv")

lgbt_index <- lgbt_index |> filter(Year== 2019)

data <- data %>%
  left_join(lgbt_index, by = c("country_name" = "Entity"))
```


```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")
  
map_data <- world %>%
  left_join(data, by = c("name" = "country_name"))

map_data$LGBT..Policy.Index <- as.numeric(as.character(map_data$LGBT..Policy.Index))

map_data <- map_data %>%
  filter(!is.na(LGBT..Policy.Index)) %>%
  group_by(admin) %>%
  slice(1) %>%  
  ungroup()

pal <- colorBin(
  palette = c("red", "orange", "yellow", "green", "blue", "purple"),
  domain = map_data$LGBT..Policy.Index,
  bins = 6,
  na.color = "transparent"
)

pmap <- leaflet(data = map_data) %>%
  addTiles() %>%
  addPolygons(
    fillColor = ~pal(LGBT..Policy.Index),
    color = "white",
    weight = 0.5,
    opacity = 1,
    fillOpacity = 0.7,
    highlight = highlightOptions(weight = 2, color = "black", bringToFront = TRUE),
    label = ~paste0(name, ": ", round(LGBT..Policy.Index, 1)),  # Entfernt das "%"
    labelOptions = labelOptions(direction = "auto")
  ) %>%
  addLegend(
    pal = pal,
    values = map_data$LGBT..Policy.Index,
    position = "topright",
    title = "LGBT Index",
    opacity = 1
  )

pmap
```
